% Based on Template from IIS, ETHz
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%                                                                 %
%%%%%     <file_name>.tex                                             %
%%%%%                                                                 %
%%%%% Author:      <author>                                           %
%%%%% Created:     <date>                                             %
%%%%% Description: <description>                                      %
%%%%%                                                                 %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter*{Acknowledgments}
I would like to express my sincere gratitude to my supervisor, Dr. Tommaso Polonelli, for his guidance and support throughout the different phases of the thesis, as well as for his continued feedback.

I also thank Prof. Dr. Luca Benini for serving as examiner and for coordinating this thesis and the corresponding assessment together with my supervisor.

I would like to acknowledge the Center for Project-Based Learning for providing a supportive environment and laboratory with the required professional tools, as well as the staff for their assistance, and my fellow students for all the insightful comments and discussions.

Furthermore, I am grateful to Prof. Simon Mitchell for encouraging me to start exploring the topic of variable-rate algorithms for dive computers.

I would also like to thank my recent diving instructors and friends with whom I have trained with for their support, discussions on dive computer functionality, algorithms and technology, and the ideas they contributed.

Finally, I thank my friends and flatmates for their support and for creating a positive environment during the duration of my studies and especially during the completion of this thesis.
